---
title: "Introduction to Bioconductor exercises"
author: "Tom Carroll"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval=F)
```


## Working with BSgenome objects using Biostrings

In these exercises will gain some experience working with the BSgenome packages.


1. Load the library BSgenome.Hsapiens.UCSC.hg19

```{r installBS, eval=TRUE,echo=FALSE}
suppressPackageStartupMessages(library(BSgenome.Hsapiens.UCSC.hg19))
```

```{r installBSs, eval=FALSE}
library(BSgenome.Hsapiens.UCSC.hg19)
```

2. Count the number of contigs.

```{r installBS2}
length(seqnames(BSgenome.Hsapiens.UCSC.hg19))
```

3. Give the sum of lengths of the 3 smallest chromosomes.

```{r installBS2dd3}
contigLengths <- seqlengths(BSgenome.Hsapiens.UCSC.hg19)
sum(contigLengths[order(contigLengths,decreasing = FALSE)][1:3])
```

4. How many unknown bases - base *N* -  are in chromosome 20

```{r installBS23}
alpFreq <-alphabetFrequency(BSgenome.Hsapiens.UCSC.hg19$chr20)
alpFreq["N"]
```

5. Extract the sequence from chromosome 20 at position 1,000,000 to 1,000,020 and retrieve the complement sequence

```{r installBS2ss3}
mySeq <- BSgenome.Hsapiens.UCSC.hg19$chr20[1000000:1000020]
complement(mySeq)
```

6. Write this complement sequence to a FASTA file.

```{r installBS2ssaa3ee, eval=FALSE}
mySet <- DNAStringSet(complement(mySeq))
names(mySet) <- "testFasta"
writeXStringSet(mySet,filepath ="Data/testFasta.fa")
```


7. Look up the position of MYC in IGV and find the genomic coordinates of its first exon.

```{r installBS2ss3ssee, eval=FALSE}
chr8:128748315-128748869
```

8. Extract the sequence for the first exon.

```{r installBS2sssss3ee}
mySeq <- BSgenome.Hsapiens.UCSC.hg19$chr8[128748315:128748869]

```

9. Compare the sequence to that found in IGV and identify start of translated region in gene 


10. Count the number of classical start codons (ATG) in the first exon.

```{r installBS2ss3aaee}
countPattern("ATG",mySeq)
```

11. Use IGV to review translation start codon for Gapdh and similarly count occurrence of ATG in exon 2 of NM_001289745 transcript. (chr12:6643976-6644027)

```{r installBS2ss3saasee}
gapdhSeq <- BSgenome.Hsapiens.UCSC.hg19$chr12[6643976:6644027]
countPattern("ATG",gapdhSeq)

```

